Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2017-02-05 acknowledge.tex
final submission for production as accepted

The authors wish to express their thanks to \myrevision{three anonymous reviewers and Editor in Chief Mariano Garc\'{i}a Fern\'{a}ndez whose insightful comments and candid suggestions greatly helped improve the original and revised manuscripts}. A message of gratitude is also extended to Art Fran\-kel at the U.S.~Geological Survey (USGS) for providing access to the Smoothed Probabilistic Seismic Hazard Analysis software and to Morgan Moschetti, also at the USGS, for providing advice on the use of the software. The authors also thank Andrzej Kijko at the University of Pretoria for providing access to the software for computing the seismicity parameters; and Mehdi Zare at the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for kindly providing the original catalog data and for his assistance in interpreting previous catalog information. Access to data provided by IIEES is also greatly appreciated. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunset Laws (SB 1510 and HB 1608, 2015--2016).

>>>>

The authors wish to express their thanks to three anonymous reviewers and Editor in Chief Mariano Garc\'{i}a Fern\'{a}ndez whose insightful comments and candid suggestions greatly helped improve the original and revised manuscripts. A message of gratitude is also extended to Art Fran\-kel at the U.S.~Geological Survey (USGS) for providing access to the Smoothed Probabilistic Seismic Hazard Analysis software and to Morgan Moschetti, also at the USGS, for providing advice on the use of the software. The authors also thank Andrzej Kijko at the University of Pretoria for providing access to the software for computing the seismicity parameters; and Mehdi Zare at the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for kindly providing the original catalog data and for his assistance in interpreting previous catalog information. Access to data provided by IIEES is also greatly appreciated. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunset Laws (SB 1510 and HB 1608, 2015--2016).

### 2017-02-05 attenuation.tex
final submission for production as accepted

The selected backbone GMPE for our analysis, as proposed by \citet{Kalkan2004}, is of the form
%
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
%
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, and $V_S$ is the shear wave velocity of the site of interest. In this equation, the distance $r$ is given by
%
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, ,
\end{equation}
%
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
%
\begin{equation}
\begin{array}{lcrlcr}
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,;
	\nonumber
\end{array}
\end{equation}
%
and $V_A$, $V_S$ and $h$ are
%
\begin{equation}
\begin{array}{lcrl}
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
%
respectively. The standard deviation of the residuals $(\sigma_{\ln y})$ expressing the random variability of the ground motions is 0.612. The value of $V_S$ is assumed to represent average surface rock sites.

@@

We should note here that even though \citet{Kalkan2004} derived the above attenuation relationship for distances $r$ up to 250 km, \citet{Zafarani2014} only used data up to 200 km in their evaluation of the various GMPEs available. Therefore, to be consistent with the latter, we set the models in the hazard analysis to compute ground motion intensities (i.e., PGAs) at distances no greater than 200 km.

>>>>

The selected backbone GMPE for our analysis, as proposed by \citet{Kalkan2004}, is of the form
%
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
%
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, and $V_S$ is the shear wave velocity of the site of interest. In this equation, the distance $r$ is given by
%
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, ,
\end{equation}
%
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. We should note here that given the fact that the smoothed seismicity method assumes all earthquakes as individual point sources, $r_{\mathit{cl}}$ is the same as the epicentral distance. According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
%
\begin{equation}
\begin{array}{lcrlcr}
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,;
	\nonumber
\end{array}
\end{equation}
%
and $V_A$, $V_S$ and $h$ are
%
\begin{equation}
\begin{array}{lcrl}
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
%
respectively. The standard deviation of the residuals $(\sigma_{\ln y})$ expressing the random variability of the ground motions is 0.612. The value of $V_S$ is assumed to represent average surface rock sites.

@@

We should note here that even though \citet{Kalkan2004} derived the above attenuation relationship for distances $r$ up to 250 km, \citet{Zafarani2014} only used data up to 200 km in their evaluation of the various GMPEs available. Therefore, to be consistent with the latter, we set the models in the hazard analysis to compute ground motion intensities (i.e., PGAs) at distances no greater than 200 km. We should also note that the values of $V_A$, $V_S$ and $h$ were kept constant throughout the calculations for this study. This was done because these are the values for which the attenuation relationship showed good agreement with PGA data from northern Iran at rock sites. While it would be tempting to adjust these values to the specific site conditions of each grid point, or to match the hypocentral depth of each earthquake in the catalog, doing this would effectively modify the goodness-of-fit of the GMPE with the data, and thus make the predictions inaccurate.

### 2017-02-05 compilation.tex
final submission for production as accepted

For the particular case of our region of interest, the catalog prepared by \citet{Zare2014} included events $M_w \geq 4$ between 550 B.C.~and 2006. We took information from the catalog up to December 1999, and augmented it by adding newer and more complete data from IIEES for events $M \geq 3$ between January 2000 and December 2015. While we recognize that earthquakes in the range $3 \leq M_w \leq 4$ are unlikely to cause structural damage, we include them as a means to feed the model with information about epicentral locations with potential for future larger magnitude earthquakes \citep{Kafka_2000_BSSA}, and to help inform the Guten\-berg-Richter relationship. By contrast, in the recent seismic hazard analysis done for Iran by \citet{Khodaverdian_2016_BSSA}, the authors used a similar catalog for events $M_w \geq 4$, but only included data through 2011, thus excluding the 2012 $M_w$ 6.4 Tabriz earthquake, which severely damaged the northeast part of the country.

>>>>

For the particular case of our region of interest, the catalog prepared by \citet{Zare2014} included events $M_w \geq 4$ between 550 B.C.~and 2006. We took information from the catalog up to December 1999, and augmented it by adding newer and more complete data from IIEES for events $M \geq 3$ between January 2000 and December 2015 (no formal catalog exists with earthquakes $M_w \leq 4$ prior to 2000). While we recognize that earthquakes in the range $3 \leq M_w \leq 4$ are unlikely to cause structural damage, we include them as a means to feed the model with information about epicentral locations with potential for future larger magnitude earthquakes \citep{Kafka_2000_BSSA}, and to help inform the Guten\-berg-Richter relationship. By contrast, in the recent seismic hazard analysis done for Iran by \citet{Khodaverdian_2016_BSSA}, the authors used a similar catalog for events $M_w \geq 4$, but only included data through 2011, thus excluding the 2012 $M_w$ 6.4 Tabriz earthquake, which severely damaged the northeast part of the country.

### 2017-02-05 conclusions.tex
final submission for production as accepted

We present results from a seismic hazard analysis for northern Iran based on a smoothed seismicity approach, including regional maps of expected peak ground accelerations for different seismic zonation models. In particular, we calculated the hazard for the region of interest using two basic models, a five-seismic zone regions model and a uniform seismic zone model, including uncertainty in the combination of both models, the seismic parameters, and the selection of a ground motion prediction equation via a logic tree approach. In general, our calculations yield results similar to those of other studies for the region of interest and the main cities in it. We, however, provide a newer and alternative perspective to the problem of seismic hazard analysis for norther Iran than previous studies because our use of a more complete seismic catalog, which includes recent damaging events and low-magnitude earthquakes. In our analysis we found that the combination of the regional and uniform models not only allowed us to account for the uncertainty in seismic zonation but it also led to a final hazard map that adjusted better to the visual interpretation of the historical and instrumental seismicity observed in the region of interest, the location of known seismically active faults, and the low-magnitude seismic activity in the region. While no one model can be deemed more appropriate than others, we believe our results constitute a positive contribution to the understanding of seismic hazard in the region, especially for rural areas that are not the subject of specialized hazard studies such as those conducted for major urban centers \myrevision{or specific sites of interest near notable faults with a history of damaging earthquakes}. Future work to \myrevision{complement and} improve the results presented here should consider other ground motion measures of intensity such as response spectral accelerations for different periods and \myrevision{site conditions, as well as alternative ground motion prediction equations suitable for northern Iran. All things considered, we recognize that additional efforts are necessary to obtain a more robust estimate of hazard, especially for specific areas of interest. In this regard, future efforts should be directed to the combination of studies like this for the contribution of the background seismicity with more targeted studies considering extended source models for areas near major fault lines. It should also be noted that this is not a final assessment of the seismic hazard of the region, and decisions regarding seismic design or risk mitigation plans should consider this as well as other studies.}

>>>>

We present results from a seismic hazard analysis for northern Iran based on a smoothed seismicity approach, including regional maps of expected peak ground accelerations for different seismic zonation models. In particular, we calculated the hazard for the region of interest using two basic models, a five-seismic zone regions model and a uniform seismic zone model, including uncertainty in the combination of both models, the seismic parameters, and the selection of a ground motion prediction equation via a logic tree approach. In general, our calculations yield results similar to those of other studies for the region of interest and the main cities in it. We, however, provide a newer and alternative perspective to the problem of seismic hazard analysis for northern Iran than previous studies because of the coverage area and our use of a more complete seismic catalog, which includes recent damaging events and low-magnitude earthquakes.

In our analysis we found that the combination of the regional and uniform models not only allowed us to account for the uncertainty in seismic zonation but it also led to a final hazard map that adjusted better to the visual interpretation of the historical and instrumental seismicity observed in the region of interest, the location of known seismically active faults, and the low-magnitude seismic activity in the region. While no one model can be deemed more appropriate than others, we believe our results constitute a positive contribution to the understanding of seismic hazard in the region, especially for rural areas that are not the subject of specialized hazard studies such as those conducted for major urban centers or specific sites of interest near notable faults with a history of damaging earthquakes.

Future work to complement and improve the results presented here should consider other ground motion measures of intensity such as response spectral accelerations for different periods and site conditions, as well as alternative ground motion prediction equations suitable for northern Iran. All things considered, we recognize that additional efforts are necessary to obtain a more robust estimate of hazard, especially for specific areas of interest. In this regard, future efforts should combine studies like this (for the contribution of the background seismicity) with more targeted studies considering extended source models (for areas near major fault lines). It should also be noted that this is not a final assessment of the seismic hazard of the region, and decisions regarding seismic design or risk mitigation plans should consider this as well as other studies.

### 2017-02-05 introduction.tex
final submission for production as accepted

Over the last three decades there have been multiple efforts to estimate the seismic parameters and seismic hazard of Iran, and improve the assessment of seismic risk in the country \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected, for instance, in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. There is, however, room for improvement. The most recent of these studies, \citet{Khodaverdian_2016_BSSA}, presents an estimation of spatially-variable seismic parameters for Iran using an approach similar to the one we use here, but does not necessarily focus on hazard. Other previous efforts have outdated catalogs, use fault-centered hazard analyses, or focus on particular locations or cities.

>>>>

Over the last three decades there have been multiple efforts to estimate the seismic parameters and seismic hazard of Iran, and improve the assessment of seismic risk in the country \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected, for instance, in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. There is, however, room for improvement. The most recent of these studies, \citet{Khodaverdian_2016_BSSA}, presents an estimation of spatially-variable seismic parameters for Iran using an approach similar to the one we use here, but does not necessarily focus on hazard, and does not include the latest available data for the region. Other previous efforts also have outdated catalogs, use fault-centered hazard analyses, or focus on particular locations or cities. By contrast, our analysis presents a regional perspective which is likely to be more useful for rural areas that are usually not the subject of specialized studies on hazardous faults.
